Large-scale text pre-training helps with dialogue act recognition, but not without fine-tuning

### 2021-03-12 reports/IWCS2021/iwcs-short-paper.tex
edits

\title{Pre-trained BERT helps with dialogue act recognition, but not very much}

@@

@@

\begin{abstract}
This document contains the instructions for preparing a manuscript for the proceedings of ACL-IJCNLP 2021.
The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
These instructions should be used for both papers submitted for review and for final versions of accepted papers.
Authors are asked to conform to all the directions reported in this document.
\end{abstract}

@@

Recently, large-scale language models, trained on massive corpora of text data, have achieved state-of-the-art results on a variety of traditional NLP tasks.
It is still uncertain whether these models might be useful for processing dialogue, and if so, whether they can adapt to make use of dialogue-specific features.

Dialogue, especially spoken dialogue, is radically different from the kind of data that neural language models are pre-trained on.
In the example from the Switchboard corpus, shown in Table~\ref{table:example}, it is evident that the syntactic structure of dialogue is very different from  written text, even if one imagines a script for a movie or a play.
Not only is the internal structure of contributions different---with features such as disfluencies, repair, incomplete sentences, and various vocal sounds---but the sequential structure of the discourse is also very different.
In dialogue, speakers take turns and switch perspectives.
More generally, the way that utterances contribute to and cohere with the discourse is different from the relationship between sentences in a piece of written text.

\begin{table}
      \small
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
    A	&sd	& Well, I'm the kind of cook that I don't normally measure things,  \\
    A	&sd	& I just kind of throw them in \\
    A	&sd	& and, you know, I don't to the point of, you know, measuring down to the exact amount that they say.  \\
    B	&sv	& That means you're a real cook. \\
    A	&bd	& \texttt{<Laughter>} Oh, is that what it means.  \\
    A	&b	& Uh-huh.  \\
    A	&x	& \texttt{<Laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw2827). Dialogue acts: \emph{sd}---Statement-non-opinion, \emph{sv}---Statement-opinion, \emph{bd}---Downplayer, \emph{b}---Backchannel, \emph{x}---Non-verbal. }
  \label{table:example}
\end{table}

We use dialogue act recognition (DAR) as a proxy task, since both the internal content of contributions, and their sequential structure have bearing on this task. DAR is the task of assigning a dialogue act tag to each utterance in the dialogue, according to dialogue act it performs.
Using this task we investigate whether transfer learning techniques based on pre-trained language models, such as BERT \citep{devlinBERTPretrainingDeep2018}, are useful for dialogue.

\paragraph{Contributions}
The main results of this paper have to do with to the usefulness of language model fine-tuning for DAR.
\begin{itemize}
  \item Standard BERT pre-training is useful for DAR, but the model performs poorly without fine-tuning (\S\ref{sec:experiment2})
  \item Further pre-training with data from the target domain shows promise for dialogue, but we find mixed results when pre-training with a larger corpus of dialogical data from outside the target domain (\S\ref{sec:experiment3}).
  \end{itemize}

In the following section we provide more details about the dialogue act recognition task and transfer learning techniques.

@@

\subsection{Dialogue Act Recognition}
The concept of a dialogue act is based on that of speech acts \citep{austinHowThingsWords2009}.
Breaking with classical semantic theory, speech act theory considers not only the propositional content of an utterance but also the actions, such as \emph{promising} or \emph{apologizing}, it carries out.
Dialogue acts extend the concept of the speech act, with a focus on the interactional nature of most speech.
DAMSL \citep{coreCodingDialogsDAMSL1997}, for example, is an influential dialogue act tagging scheme where dialogue acts are defined in part by whether they have a \emph{forward-looking} function (are expecting a response) or \emph{backward-looking} function (are in response to a previous utterance).

DAR is the task of labeling utterances with the dialogue act they perform from a given set of dialogue act tags.
As with other sequence labeling tasks in NLP, some notion of context is helpful in DAR.
One of the first performant machine learning models for DAR was a Hidden Markov Model that used various lexical and prosodic features as input \citep{stolckeDialogueActModeling2000}.

Most successful neural approaches also model some notion of context.
We follow previous work that models dialogue act sequences with an RNN over utterances \citep{kalchbrennerRecurrentConvolutionalNeural2013,tranHierarchicalNeuralModel2017,botheContextbasedApproachDialogue2018,bothe2018conversational,zhao2018unified}.
Our model predicts the most likely tag given the hidden state of the RNN at each position, though other work has shown that DAR benefits from more sophisticated approaches, such as conditional random field \citep{chenDialogueActRecognition2017} or uncertainty propagation \citep{tranPreservingDistributionalInformation2017}.

@@

Recently, neural models trained on large datasets with various unsupervised tasks have found success creating sentence-level and contextual word representations. 
BERT, a multi-layer transformer, is one such model \citep{devlinBERTPretrainingDeep2018}. 
BERT is pre-trained on two unsupervised tasks: \emph{masked token prediction} and \emph{next sentence prediction}.
In masked token prediction, some percentage of words are randomly replaced with a mask token, and the model is trained to predict the identity of the original token, based on the context sentence.
In next sentence prediction, the model is given two sentences and trained to predict whether the second sentence follows the first in the original text, or if it was randomly chosen from elsewhere in the corpus.
After pre-training, BERT can be applied to a supervised task by adding additional un-trained layers that take the hidden state of one or more of BERT's layers as input.

There is some previous work applying BERT to dialogue.
\citet{baoPLATOPretrainedDialogue2019} and \citet{chenSemanticallyConditionedDialog2019a} both use BERT for dialogue generation tasks.
Similarly, \citet{vigComparisonTransferLearningApproaches2019} find BERT useful for selecting a response from a list of candidate responses in a dialogue.
\citet{mehriPretrainingMethodsDialog2019} evaluate BERT in various dialogue tasks including DAR, and find that a model incorporating BERT outperforms a baseline model.
Finally, \citet{chakravarty2019dialog} use BERT for dialogue act classification for a proprietary domain and achieves promising results, and \citet{ribeiro2019deep} surpass the previous state-of-the-art on generic dialog act recognition for Switchboard and MRDA corpora. In our work we are not attempting to beat the current state-of-the-art due to the different aims of the study.

@@

\citet{petersTuneNotTune2019} investigate when it is best to fine-tune BERT for sentence classification tasks.
They find that when the target task is very similar to the pre-training task, fine-tuning provides less of a performance boost.
We note that there is some conceptual relationship between DAR and next sentence prediction, since the dialogue act constrains (or at least is predictive of) the dialogue act that follows it.

\citet{petersTuneNotTune2019} also consider whether the similarity of the pre-training and target domains may have similar consequences, but don't find an effect of domain similarly for the natural language inference tasks they tested on.
As discussed in \S\ref{sec:data}, however, there are dramatic differences between written text and transcribed dialogue which may make fine-tuning necessary for DAR.

\paragraph{} In the next section, we will present the datasets that we will use in our experiments along with their distinctive features.

\section{Data}\label{sec:data}
We perform experiments on the Switchboard Dialogue Act Corpus (SWDA), which is a subset of the larger Switchboard corpus, and the dialogue act-tagged portion of the AMI Meeting Corpus (AMI-DA). 
SWDA is tagged with a set of 220 dialogue act tags which, following  \citet{jurafskySwitchboardSWBDDAMSLShallowDiscourseFunction1997a}, we cluster into a smaller set of 42 tags.
AMI uses a smaller tagset of 16 dialogue acts \citep{GuidelinesDialogueAct2005}.

@@

\begin{table}[]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Switchboard}       & \textbf{AMI Corpus}                     \\ \midrule
Dyadic                     & Multi-party                             \\
Casual conversation        & Mock business meeting                   \\
Telephone                  & In-person \& video                      \\ \midrule
English                    & English                                 \\ 
Native speakers            & Native \& non-native speakers           \\ 
early '90s                 & 2000s                                   \\ \midrule
2200 conversations         & 171 meetings                            \\
  \hspace{1em} 1155 in SWDA               & \hspace{1em} 139 in AMI-DA                           \\
400k utterances             & 118k utterances                         \\
3M tokens                  & 1.2M tokens                             \\ \bottomrule
\end{tabular}
  \caption{Comparison between Switchboard and the AMI Meeting Corpus}
  \label{table:corpora}
\end{table}

@@

Utterances are tokenized using with a word piece tokenizer \citep{Wu2016} with a vocabulary of 30,000.
We also add five speaker tokens and prepend each utterance with a speaker token that uniquely identifies the corresponding speaker within that dialogue.

\paragraph{Dialogue act grouping}
For our corpora we also considered grouping the dialogue acts according to DAMSL coders manual \citep{jurafskySwitchboardSWBDDAMSLShallowDiscourseFunction1997a}: Forward-Communicative-Function (51.86\% of all DAs), Backwards-Communicative-Function (30.35\%), Other (9.04\%) and Communicative-Status (8.75\%).
This allows the direct comparison between SWDA and AMI-DA.
With respect to laughs, we observed an even distribution of them in the groups, except for the Communicative-Status group in the case of SWDA, because of the ``Nonverbal'' dialogue acts (see footnote~1).

@@

We also experiment with three unlabeled dialogue corpora (Table~\ref{tab:pretraining-corpora}), which we use to provide further pre-training for the BERT encoder.

@@

We also experiment with a much larger a corpus constructed from OpenSubtitles \citep{Lison2016}.
Because utterances are not labeled with speaker in the OpenSubtitles corpus, we randomly assigned a speaker token to each utterance to maintain the format of the other dialouge corpora.

@@

To test the effectiveness of BERT for DAR, we employ a simple neural architecture with two components: an encoder that vectorizes utterances, and a sequence model that predicts dialogue act tags from the vectorized utterances (Figure~\ref{fig:model-architecture}).
Since we are primarily interested in comparing different utterance encoders, we use a basic RNN as the sequence model in every configuration.\footnote{We have experimented with LSTM as the sequence model, but the accuracy was not significantly different compared to RNN. It can be explained by the absence of longer distance dependencies on this level of our model.} 
The RNN takes the encoded utterance as input at each time step,
and its hidden state is passed to a simple linear classification layer over dialogue act tags.
Conceptually, the encoded utterance represents the context-agnostic features of the utterance, and the hidden state of the RNN represents the full discourse context.

\begin{figure*}
  \include{model}
  \caption{Simple neural dialogue act recognition sequence model}
  \label{fig:model-architecture}
\end{figure*}

As a baseline utterance encoder, we use a word-level CNN with window sizes of 3, 4, and 5, each with 100 feature maps \citep{kimConvolutionalNeuralNetworks2014}. 
The model uses 100-dimensional word embeddings, which are initialized with pre-trained gloVe vectors \citep{penningtonGloveGlobalVectors2014}.

For the BERT utterance encoder, we use the BERT\textsubscript{BASE} model with hidden size of 768 and 12 transformer layers and self-attention heads \citep[][\S3.1]{devlinBERTPretrainingDeep2018}.
In our implementation, we use the un-cased model provided by \citet{wolfHuggingFaceTransformersStateoftheart2019}.

The next section will introduce the experiments that we have carried out and will discuss the results that we obtained from them.

\subsection{Experiment 2: Impact of pre-training vs. fine-tuning} \label{sec:experiment2} % Bill
Next, we analyze how pre-training affects BERT's performance as an utterance encoder.
To do so, we consider the performance of DAR models with three different utterance encoders:
\begin{itemize}
  \item \texttt{BERT-FT} -- pre-trained BERT with DAR fine-tuning 
  \item \texttt{BERT-RI} -- randomly initialized BERT (with DAR fine-tuning)
  \item \texttt{BERT-FZ} -- pre-trained BERT without fine-tuning (frozen during DAR training)
\end{itemize}

\begin{table}[]
\centering
\begin{tabular}{@{}lrrrr@{}}
\toprule
                           & \multicolumn{2}{c}{SWDA}                          & \multicolumn{2}{c}{AMI-DA}                        \\ \midrule
                           & \multicolumn{1}{c}{F1} & \multicolumn{1}{c}{acc.} & \multicolumn{1}{c}{F1} & \multicolumn{1}{c}{acc.} \\
\texttt{BERT-FT} & 36.75           & 76.60                    & 43.42                  & 64.93                    \\
\texttt{BERT-RI} & 32.18           & 73.80                    & 34.88                  & 60.89                    \\
\texttt{BERT-FZ} & 7.75            & 55.61                    & 14.86                  & 48.34 \\ \midrule
    Majority class    & 0.78  & 33.56 &  1.88 & 28.27      \\ \bottomrule
\end{tabular}
  \caption{Macro-average F1 and accuracy of BERT with standard pre-training and DAR fine-tuning (\texttt{BERT-FT}) vs.~the same model without pre-training (\texttt{BERT-RI}) and without fine-tuning (\texttt{BERT-FZ}).}
  \label{table:exp2}
\end{table}

The pre-trained model is more accurate than the randomly initialized model by several percentage points on both DA corpora,
suggesting that BERT's extensive pre-training does provide some useful information for recognizing different dialogue acts (Table~\ref{table:exp2}).
This performance boost is much more pronounced in the macro-averaged F1 score,
which is explained by the fact that at the tag level, pre-training has a larger impact on less frequent tags.
For a disaggregated view, see Figure \ref{fig:f1-by-da} in Appendix~\ref{sec:suppl}.

@@

Comparing the performance of the fine-tuned encoder to that of the frozen model, we note that pre-training alone is not enough.
The frozen model is heavily biased towards the most frequent tags, which explains its especially poor macro-average score (Table~\ref{table:exp2}).
In SWDA, for example, the model with a frozen encoder predicts one of the two most common tags (Statement-non-opinion or Acknowledge) 86\% of the time, whereas those two
tags account for only 51\% of the ground truth tags.
The fine-tuned model is much less biased; it predicts the two most common tags only 59\% of the time.

@@

\begin{table}[]
\begin{tabular}{@{}lrrrr@{}}
\toprule
           & \multicolumn{2}{c}{SWDA}                          & \multicolumn{2}{c}{AMI-DA}                        \\ \midrule
           & \multicolumn{1}{c}{F1} & \multicolumn{1}{c}{acc.} & \multicolumn{1}{c}{F1} & \multicolumn{1}{c}{acc.} \\
\texttt{BERT-FT}    & 36.75                  & 76.60                    & 43.42                  & 64.93                    \\
\texttt{BERT+ID-FT} & 43.63                  & 77.01                    & 46.70                  & \textbf{68.88}           \\
\texttt{BERT+CC-FT} & \textbf{47.78}         & \textbf{77.35}           & \textbf{48.86}         & 68.79                    \\ 
\texttt{BERT+OS-FT} & 41.42                  & 76.95                    & 48.65                  & 68.07                    \\ \midrule
\texttt{BERT-FZ}    &  7.75                  & 55.61           & 14.86                  & 48.34                    \\
\texttt{BERT+ID-FZ} &  6.46                  & 52.30                    & 14.48                  & 48.18                    \\
\texttt{BERT+CC-FZ} &  5.76                  & 51.14                    & 11.34                  & 40.48                    \\ 
\texttt{BERT+OS-FZ} &  \textbf{9.60}         & \textbf{57.67}           & \textbf{17.03}         & \textbf{51.03}           \\ \midrule
    Majority class  & 0.78                   & 33.56                    &  1.88                  & 28.27      \\ \bottomrule
\end{tabular}
\caption{Comparison of macro-average F1 and accuracy with further in-domain (\texttt{ID}), combined corpus (\texttt{CC}), and OpenSubtitles (\texttt{OS}) dialogue pre-training,
    for the frozen (\texttt{FZ}) and fine-tuned (\texttt{FT}) conditions.}
  \label{table:exp3}
\end{table}

@@

\bibliography{acl2020}
\bibliographystyle{acl_natbib}
\appendix
\onecolumn
\bibliographystyle{acl_natbib}
\bibliography{iwcs-short-paper}

>>>>

\title{Large-scale text pre-training helps with dialogue act recognition, but not without fine-tuning}

@@

% We should make the comparison with other tasks where BERT representations work well AS IS (i.e., without finetuning) vs dialogue where 
% the pretraining is helpful, but not without finetuning
% Missing reference(s) from *SEM reviews
% Usefulness of in-domain pre-training for spoken dialogue
% Future work: dialogue act recognition is not enough to fully evaluate BERT for dialogue - other turn change prediction, intent classification, utterance segmentation and disfluency detection
%

@@

\begin{abstract}
  We use dialogue act recognition (DAR)
  to investigate how well BERT represents utterances in dialogue,
  and how fine-tuning and large-scale pre-training contribute to its performance.
  We find that while both the standard BERT pre-training and pretraining on
  dialogue-like data are useful, task-specific fine-tuning is essential for good performance.
\end{abstract}

@@

Recently, large-scale neural language models, trained on massive corpora of text data, 
have achieved state-of-the-art results on a variety of traditional NLP tasks.
Given that dialogue, especially spoken dialogue, 
is radically different from the kind of data these language models are pre-trained on,
it is uncertain whether these models would be useful for dialouge-oriented tasks.
In the example from the Switchboard corpus, shown in Table~\ref{tab:example}, 
it is evident that the structure of dialogue is quite different from that of written text.
%even if one imagines a script for a movie or a play.
Not only is the internal structure of the contributions different---%
with features such as disfluencies, repair, incomplete sentences, and various vocal sounds---%
but the sequential structure of the discourse is different as well.
In dialogue, speakers take turns and switch perspectives.
More generally, the way that utterances contribute to and cohere with the discourse
is different from the relationship between sentences in a piece of written text.

In this paper, we investigate how well 
one such large-scale language model,
BERT \citep{devlinBERTPretrainingDeep2018}, 
represents utterances in dialogue.
We use dialogue act recognition (DAR) as a proxy task, 
since both the internal content and the sequential structure of utterances 
has bearing on this task

\begin{table}
      \small
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
    A	&sd	& Well, I'm the kind of cook that I don't normally measure things,  \\
    A	&sd	& I just kind of throw them in \\
    A	&sd	& and, you know, I don't to the point of, you know, measuring down to the exact amount that they say.  \\
    B	&sv	& That means you're a real cook. \\
    A	&bd	& \texttt{<Laughter>} Oh, is that what it means.  \\
    A	&b	& Uh-huh.  \\
    A	&x	& \texttt{<Laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw2827). Dialogue acts: \emph{sd}---Statement-non-opinion, \emph{sv}---Statement-opinion, \emph{bd}---Downplayer, \emph{b}---Backchannel, \emph{x}---Non-verbal. }
  \label{tab:example}
\end{table}

We have two main contributions.
First we find that while standard BERT pre-training is useful,
but the model performs poorly without fine-tuning (\S\ref{sec:experiment2}).
Second, we find that further pre-training with data from the target domain shows promise for dialogue, 
but the results are mixed when pre-training with a larger corpus of dialogical data from outside the target domain (\S\ref{sec:experiment3}).

@@

\subsection{Dialogue Act Recognition}
The concept of a dialogue act is based on that of speech acts \citep{austinHowThingsWords2009}.
Breaking with classical semantic theory, speech act theory considers not only the propositional content of an utterance but also the actions, such as \emph{promising} or \emph{apologizing}, it carries out.
Dialogue acts extend the concept of the speech act, 
with a focus on the interactional nature of most speech.
%DAMSL \citep{coreCodingDialogsDAMSL1997} is an influential dialogue act tagging scheme 
%where dialogue acts are defined in part by whether they have a 
%\emph{forward-looking} function (are expecting a response) or 
%\emph{backward-looking} function (are in response to a previous utterance).

DAR is the task of labeling utterances with the dialogue act they perform 
from a given set of dialogue act tags.
As with other sequence labeling tasks in NLP, some notion of context is helpful in DAR.
One of the first performant machine learning models for DAR was a Hidden Markov Model 
that used various lexical and prosodic features as input \citep{stolckeDialogueActModeling2000}.
Most successful neural approaches also model some notion of context \citep[e.g.,][]{kalchbrennerRecurrentConvolutionalNeural2013,tranHierarchicalNeuralModel2017,botheContextbasedApproachDialogue2018,bothe2018conversational,zhao2018unified}.
Following previous work, we model dialogue act sequences with an RNN over utterance representations,
which are constructed by BERT.

@@

%Recently, neural models trained on large datasets with various unsupervised tasks have found success creating sentence-level and contextual word representations. 
BERT, a multi-layer transformer model \citep{devlinBERTPretrainingDeep2018}, 
is pre-trained on two unsupervised tasks: 
\emph{masked token prediction} and \emph{next sentence prediction}.
In masked token prediction, some percentage of words are randomly replaced with a mask token.
The model is trained to predict the identity of the original token based on the context sentence.
In next sentence prediction, the model is given two sentences and trained to predict whether the second sentence follows the first in the original text or if it was randomly chosen from elsewhere in the corpus.
After pre-training, BERT can be applied to a supervised task by adding additional un-trained layers that take the hidden state of one or more of BERT's layers as input.

There is some previous work applying BERT to dialogue.
\citet{baoPLATOPretrainedDialogue2019} and \citet{chenSemanticallyConditionedDialog2019a} both use BERT for dialogue generation tasks.
Similarly, \citet{vigComparisonTransferLearningApproaches2019} find BERT useful for selecting a response from a list of candidate responses in a dialogue.
\citet{mehriPretrainingMethodsDialog2019} evaluate BERT in various dialogue tasks including DAR, and find that a model incorporating BERT outperforms a baseline model.
Finally, \citet{chakravarty2019dialog} use BERT for dialogue act classification for a proprietary domain and achieves promising results, and \citet{ribeiro2019deep} surpass the previous state-of-the-art on generic dialog act recognition for Switchboard and MRDA corpora. 
This paper aims to supplement the findings of previous work by investigating how much of 
BERT's sucess for dialogue tasks is due to its extensive pre-training and how much is due to 
task-specific fine-tuning.

@@

\citet{petersTuneNotTune2019} investigate when it is best to fine-tune BERT for sentence classification tasks 
and find that when the target task is very similar to the pre-training task, 
fine-tuning provides less of a performance boost.
We note that there is some conceptual relationship between DAR and next sentence prediction, since the dialogue act constrains (or at least is predictive of) the dialogue act that follows it.
That said, the discourse strucutre of the encyclopedia and book data that makes up BERT's pre-training corpus
is probably quite different from that of natural dialogue.
%\citet{petersTuneNotTune2019} also consider whether the similarity of the pre-training and target domains may have similar consequences, but don't find an effect of domain similarly for the natural language inference tasks they tested on.
%As discussed in \S\ref{sec:data}, however, there are dramatic differences between written text and transcribed dialogue which may make fine-tuning necessary for DAR.

\section{Data}\label{sec:data}

\begin{table}[]
\centering
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Switchboard}       & \textbf{AMI Corpus}                     \\ \midrule
Dyadic                     & Multi-party                             \\
Casual conversation        & Mock business meeting                   \\
Telephone                  & In-person \& video                      \\ \midrule
English                    & English                                 \\ 
Native speakers            & Native \& non-native speakers           \\ 
early '90s                 & 2000s                                   \\ \midrule
2200 conversations         & 171 meetings                            \\
  \hspace{1em} 1155 in SWDA               & \hspace{1em} 139 in AMI-DA                           \\
400k utterances             & 118k utterances                         \\
3M tokens                  & 1.2M tokens                             \\ \bottomrule
\end{tabular}
  \caption{Comparison between Switchboard and the AMI Meeting Corpus}
  \label{tab:corpora}
\end{table}

We perform experiments on the Switchboard Dialogue Act Corpus (SWDA), 
which is a subset of the larger Switchboard corpus, 
and the dialogue act-tagged portion of the AMI Meeting Corpus (AMI-DA). 
SWDA is tagged with a set of 220 dialogue act tags which, following  \citet{jurafskySwitchboardSWBDDAMSLShallowDiscourseFunction1997a}, we cluster into a smaller set of 42 tags.
AMI uses a smaller tagset of 16 dialogue acts \citep{GuidelinesDialogueAct2005}.
See Table~\ref{tab:corpora} for details.

@@

@@

Utterances are tokenized with BERT's word piece tokenizer with a vocabulary of 30,000.
To this vocabulary we added five speaker tokens and prepend each utterance with a speaker 
token that uniquely identifies the corresponding speaker within that dialogue.

%\paragraph{Dialogue act grouping}
%For our corpora we also considered grouping the dialogue acts according to DAMSL coders manual \citep{jurafskySwitchboardSWBDDAMSLShallowDiscourseFunction1997a}: Forward-Communicative-Function (51.86\% of all DAs), Backwards-Communicative-Function (30.35\%), Other (9.04\%) and Communicative-Status (8.75\%).
%This allows the direct comparison between SWDA and AMI-DA.
%With respect to laughs, we observed an even distribution of them in the groups, except for the Communicative-Status group in the case of SWDA, because of the ``Nonverbal'' dialogue acts (see footnote~1).

@@

We also experiment with three unlabeled dialogue corpora, which we use to provide further pre-training for the BERT encoder.

@@

We also experiment with a much larger a corpus (350M tokens) constructed from OpenSubtitles \citep{Lison2016}.
Because utterances are not labeled with speaker, we randomly assigned a speaker token to each utterance to maintain the format of the other dialouge corpora.

@@

We use a simple neural architecture with two components: 
an encoder that vectorizes utterances (BERT), 
and a sequence model that takes the utterance representations as input.
Since we are primarily interested in comparing different training and fine-tuning schemes for the utterance encoder, 
we use a basic single-layer RNN as the sequence model in every configuration.\footnote{We have experimented with LSTM as the sequence model, but the accuracy was not significantly different compared to RNN. It can be explained by the absence of longer distance dependencies on this level of our model.}
At each time step, the RNN takes the encoded utterance as input 
and its hidden state is passed to a simple linear classification layer over dialogue act tags.\footnote{
Other work has shown that DAR benefits from more sophisticated decoding, such as conditional random field \citep{chenDialogueActRecognition2017} and uncertainty propagation \citep{tranPreservingDistributionalInformation2017}.}

Conceptually, the encoded utterance represents the context-agnostic features of the utterance 
and the hidden state of the RNN represents the full discourse context.

%As a baseline utterance encoder, we use a word-level CNN with window sizes of 3, 4, and 5, each with 100 feature maps \citep{kimConvolutionalNeuralNetworks2014}. 
%The model uses 100-dimensional word embeddings, which are initialized with pre-trained gloVe vectors \citep{penningtonGloveGlobalVectors2014}.

For the BERT utterance encoder, we use the BERT\textsubscript{BASE} model with hidden size of 768 and 12 transformer layers and self-attention heads \citep[][\S3.1]{devlinBERTPretrainingDeep2018}.
In our implementation, we use the un-cased model provided by \citet{wolfHuggingFaceTransformersStateoftheart2019}.
The RNN has a hidden layer size of 100.

\subsection{Pre-training vs. fine-tuning} \label{sec:experiment2} % Bill
First, we analyze how pre-training affects BERT's performance as an utterance encoder.
To do so, we consider the performance of DAR models with three different utterance encoders:
\begin{itemize}
  \item \texttt{BERT-FT} -- pre-trained BERT with DAR fine-tuning 
  \item \texttt{BERT-RI} -- randomly initialized BERT (with DAR fine-tuning)
  \item \texttt{BERT-FZ} -- pre-trained BERT without fine-tuning (frozen during DAR training)
\end{itemize}

\begin{table}[]
\centering
\begin{tabular}{@{}lrrrr@{}}
\toprule
                           & \multicolumn{2}{c}{SWDA}                          & \multicolumn{2}{c}{AMI-DA}                        \\ \midrule
                           & \multicolumn{1}{c}{F1} & \multicolumn{1}{c}{acc.} & \multicolumn{1}{c}{F1} & \multicolumn{1}{c}{acc.} \\
\texttt{BERT-FT} & 36.75           & 76.60                    & 43.42                  & 64.93                    \\
\texttt{BERT-RI} & 32.18           & 73.80                    & 34.88                  & 60.89                    \\
\texttt{BERT-FZ} & 7.75            & 55.61                    & 14.86                  & 48.34 \\ \midrule
    Majority class    & 0.78  & 33.56 &  1.88 & 28.27      \\ \bottomrule
\end{tabular}
  \caption{Macro-average F1 and accuracy of BERT with standard pre-training and DAR fine-tuning (\texttt{BERT-FT}) vs.~the same model without pre-training (\texttt{BERT-RI}) and without fine-tuning (\texttt{BERT-FZ}).}
  \label{tab:exp2}
\end{table}

The pre-trained model is more accurate than the randomly initialized model by several percentage points on both DA corpora,
suggesting that BERT's extensive pre-training does provide some useful information for recognizing different dialogue acts (Table~\ref{tab:exp2}).
This performance boost is much more pronounced in the macro-averaged F1 score,
which is explained by the fact that at the tag level, pre-training has a larger impact on less frequent tags.
For a disaggregated view, see Figure \ref{fig:f1-by-da} in Appendix~\ref{sec:suppl}.

@@

Comparing the performance of the fine-tuned encoder to that of the frozen model, we note that pre-training alone is not enough.
The frozen model is heavily biased towards the most frequent tags, which explains its especially poor macro-average score (Table~\ref{tab:exp2}).
In SWDA, for example, the model with a frozen encoder predicts one of the two most common tags (Statement-non-opinion or Acknowledge) 86\% of the time, whereas those two
tags account for only 51\% of the ground truth tags.
The fine-tuned model is much less biased; it predicts the two most common tags only 59\% of the time.

@@

\begin{table}[]
\begin{tabular}{@{}lrrrr@{}}
\toprule
           & \multicolumn{2}{c}{SWDA}                          & \multicolumn{2}{c}{AMI-DA}                        \\ \midrule
           & \multicolumn{1}{c}{F1} & \multicolumn{1}{c}{acc.} & \multicolumn{1}{c}{F1} & \multicolumn{1}{c}{acc.} \\
\texttt{BERT-FT}    & 36.75                  & 76.60                    & 43.42                  & 64.93                    \\
\texttt{BERT+ID-FT} & 43.63                  & 77.01                    & 46.70                  & \textbf{68.88}           \\
\texttt{BERT+CC-FT} & \textbf{47.78}         & \textbf{77.35}           & \textbf{48.86}         & 68.79                    \\ 
\texttt{BERT+OS-FT} & 41.42                  & 76.95                    & 48.65                  & 68.07                    \\ \midrule
\texttt{BERT-FZ}    &  7.75                  & 55.61           & 14.86                  & 48.34                    \\
\texttt{BERT+ID-FZ} &  6.46                  & 52.30                    & 14.48                  & 48.18                    \\
\texttt{BERT+CC-FZ} &  5.76                  & 51.14                    & 11.34                  & 40.48                    \\ 
\texttt{BERT+OS-FZ} &  \textbf{9.60}         & \textbf{57.67}           & \textbf{17.03}         & \textbf{51.03}           \\ \midrule
    Majority class  & 0.78                   & 33.56                    &  1.88                  & 28.27      \\ \bottomrule
\end{tabular}
\caption{Comparison of macro-average F1 and accuracy with further in-domain (\texttt{ID}), combined corpus (\texttt{CC}), and OpenSubtitles (\texttt{OS}) dialogue pre-training,
    for the frozen (\texttt{FZ}) and fine-tuned (\texttt{FT}) conditions.}
  \label{tab:exp3}
\end{table}

@@

\bibliographystyle{acl_natbib}
\bibliography{acl2020}
